Reproducible Research Report
============================================

# Reproducible Research: Peer Assessment 1
This ist the report of my approach how I processed and analyzed the data for the first peer assignment of COursera course 'Reproducible Research'.

## Loading and processing the data 

The used is available under the following [link] (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip). After the data is loaded you need to unzip the archive file and load the contained csv-file 'activity.csv'

```{r}
library(data.table)
act<-fread("activity.csv")
dim(act)
summary(act)
head(act)
```

## What is mean total number of steps taken per day?

1 Make a histogram of the total number of steps taken each day
```{r}
t_st<-tapply(act$steps,act$date,FUN=function(x) sum(x,na.rm=TRUE))
plot(t_st,col="red",type="S",xaxt="n",bg="blue",main="Total number of steps",
                xlab="Date",ylab="Totals")
axis(1, at=seq(1, 61, by=1),labels=names(t_st))
day_totals<-data.frame(day=names(t_st),total_steps=t_st)
```

2 Calculate and report the mean and median total number of steps taken per day
```{r}
mean_steps<-mean(day_totals$total_steps,na.rm=TRUE)
median_steps<-median(day_totals$total_steps,na.rm=TRUE)
```

The mean total number of steps taken per day is `r mean_steps` and the median is equal to `r median_steps`.


## What is the average daily activity pattern?

1 Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
m_st_int<-act[,mean(steps,na.rm=TRUE),by=act$interval]
plot(m_st_int$act,m_st_int$V1,main="Average number of steps taken per interval",xlab="Intervals",ylab="Mean values per interval",type="n")
lines(m_st_int$act,m_st_int$V1,col="blue")
```


2    Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
m_max_st<-m_st_int[max(m_st_int$V1),"act",with=FALSE]
```
Interval `r m_max_st` contains the maximum number of steps

## Imputing missing values

1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
##get the vector for complete.cases
cc<-complete.cases(act$steps,act$date,act$interval)
##number of rows with missing values
nr_cc<-length(cc[!cc])
```

Total number of rows with missing values is `r nr_cc`